---
title: 'A04: Ethics and Privacy'
author: "Cadence Fisher"
date: '2022-04-16'
slug: a04-ethics-and-privacy
categories: []
tags: []
---
This assignment consists of three parts.
# Part 1: Dealing with Misleading Data
## Creating the Data Frame
```{r}
library(tidyverse)
```

```{r}
datamask <- tribble(
  ~mask, ~nomask, ~day,
  25, 10, 1,
  19.5, 9, 2,
  19.5, 9.2, 3,
  20.2, 9.8, 4,
  19.6, 9.9, 5,
  19.6, 9.2, 6,
  20.2, 9.2, 7,
  19.8, 8.9, 8,
  20.4, 8.4, 9,
  21.5, 8.4, 10,
  19.8, 8.6, 11,
  19.8, 8.5, 12,
  20.4, 9.8, 13,
  19, 9.9, 14,
  19.5, 10.1, 15,
  17, 9.5, 16,
  16.3, 9.5, 17,
  16.5, 9.5, 18,
  16.6, 10, 19,
  16, 8.7, 20,
  16.2, 9, 21,
  15.9, 8.8, 22,
  16, 9, 23
)
```
```{r}
attach(datamask)
```


## Make a more accurate visualization of the Data
```{r}
ggplot(datamask) +
  geom_line(aes(day, mask), color = "green") +
  geom_line(aes(day, nomask), color = "blue") +
  labs(
    x = 'Date (Beginning 7/12/2020 and Ending 8/3/2020)',
    color = "Mask or No Mask",
    y = "Prevalence of COVID per 100K population",
    title = "Kansas COVID-19 7-day rolling average of Daily Cases/Per100K Population ",
    subtitle = "Mask mandated counties (green) vs. no mask mandate counties (blue)"
  ) 
```


## What Message is more clear in my visualization?
The message of how masks are impacting the liklihood of getting covid is more clear in this visualization. The original graph made it look like masks made the prevalence of covid drop overtime, to the point where cases were lower than that of non mask wearers, however this graph with an accurately scaled Y for both line plots shows that people who wore no mask were actaully far lower overall. 

## What Useful information does this data and my visualization tell about mask wearing and covid
My graph indicates that a mask mandate did not seem to decrease covid rates per 100K of the population very significantly. This also indicates that the counties without the mask mandates had low values the whole time, far lower than the counties that did have a mask mandate. This could suggest that there are other factors at play that are making the mask-mandated counties have higher overall values, and that while masks did decrease their rate of covid in those counties, counties without mask mandates still had far lower numbers. 

# Review of "How I'm Fighting Bias in Algorithms"
In this video, Joy Boulamwini discusses something called  'the coded gaze', which is her term for algorithmic bias. She gave an example of when she went to a conference where a startup had a robot with facial recognition software, and as they were giving a demo, the robot could recognize everyone but her. Not only does this show that this algorithm clearly had a very specific bias against people with darker skin, but it also made the startup (in my opinion) seem very unprofessional. It turned out that they were using a generic facial recognition software. This generic facial recognition software that multitudes of programmers use as a base code has a incredibly strong bias against people of a darker skin tone because when the people originally wrote it, they failed to take into account people with darker skin, and made their algorithm tailored to recognizing lighter skin. This is particularly a problem because one faulty code leads to algorithmic bias being spread all around the world. If training sets for algorithms are not diverse enough, this is how these biases can be created. She proposes that being careful to create adequately diverse training sets can make an algorithm far better, and have way less bias. This could be essential as police departments and other crime-fighting organizations are beginning to rely upon facial recognition. It would be horrific if the software falsely recognized someone as a criminal, or if it did not recognize one and a criminal got away. This is just one example of an algorithmic bias spreading into the real world to create real bias in people's minds as well. I felt like this video was very enlightening about the real world consequences of having a biased algorithm, and it made me understand the universal impact that code and algorithms can have on the human world and how people look at things. 


# Imbeding picture of certificate

![Alt text](/Users/cfish/Downloads/privacy_certificate_datacamp)


