---
title: 'A03 Logistic regression '
author: "Cadence Fisher"
date: '2022-04-02'
slug: a03-logistic-regression
categories: []
tags: []
---
Use the Titanic passenger data (name, age, price of ticket, etc) to try to predict who will survive and who will die

your submission should have:

a "PassengerId" column containing the IDs of each passenger from test.csv.
a "Survived" column (that you will create!) with a "1" for the rows where you think the passenger survived, and a "0" where you predict that the passenger died.
#Loading in / Preparing the data
```{r}
library(openintro)
library(tidyverse)

```
```{r}
train <- read_csv("https://raw.githubusercontent.com/fisherca/website_experiment/main/train.csv")
attach(train)
glimpse(train)
```
```{r}
test <- read_csv("https://raw.githubusercontent.com/fisherca/website_experiment/main/test.csv")
attach(test)
glimpse(test)
```


# Experimenting with the Data
## Quick tables to explore patterns
This was not in the instructions, but I thought it would help me visualize the data
```{r}
table(train$Survived, train$Sex)
```
More Females survived than males, and a far higher proportion of females survived overall
```{r}
table(train$Survived, train$Pclass)
```
```{r}
136/(80+136)
87/(97+87)
119/(372+119)
```
62% of the First class passengers survived, wheras only 47% of second class and 24% of third class passengers survived

```{r}
table(train$Survived, train$SibSp)
```
```{r}
210/(398+210)
(112+13+4+3+1)/(97+112+15+13+12+4+15+3+5+1+7)
```

34.5% of people who had no siblings or spouse survived, whereas 46% of people with at least one sibling or spouse survived

```{r}
table(train$Survived, train$Parch)
```
```{r}
233/(445+233)
(65+40+3+1+1)/(65+40+3+1+1+53+40+2+4+4+1)
```

Only 34% of people with no parents or children survived, but 51% of people with at least 1 parent or child survived



## Is gender a good predictor of death? Recreation of exercise done in tutorial
```{r}
women_tot <- train %>%
  filter(Sex == 'female')
women_alive <- train %>%
  filter(Sex == 'female', Survived == 1)
women_rate = count(women_alive)/count(women_tot)
women_rate
```

74% of the women in the train data survived
```{r}
men_tot <- train %>%
  filter(Sex == 'male')
men_alive <- train %>%
  filter(Sex == 'male', Survived == 1)
men_rate = count(men_alive)/count(men_tot)
men_rate
```
18% of the men in the train data survived

#Building a random forest model:
We are going to look at Pclass,Age, Fare Sex, SibSp, and Parch as potential predictors for Survived
##Preparing the Data
adding empty row to test data
```{r}
test$Survived <- NA
```
removing rows I will not use
```{r}
train <- train[c("PassengerId", "Survived", "Pclass", "Sex", "Age", "SibSp", "Parch", "Fare")]

test <- test[c("PassengerId", "Survived", "Pclass", "Sex", "Age", "SibSp", "Parch", "Fare")]
```

replacing NA values and changing data types if needed 
```{r}
train[is.na(train$Age), "Age"] <- median(train$Age, na.rm = TRUE)

test[is.na(test$Age), "Age"] <- median(test$Age, na.rm = TRUE)
```
```{r}
train[is.na(train$Fare), "Fare"] <- median(train$Fare, na.rm = TRUE)

test[is.na(test$Fare), "Fare"] <- median(test$Fare, na.rm = TRUE)
```
```{r}
train <- transform(train, Survived = as.factor(Survived), Pclass = as.factor(Pclass), Sex = as.factor(Sex))

test <- transform(test, Pclass = as.factor(Pclass), Sex = as.factor(Sex))
```

making a full dataset
```{r}
full_data <- rbind(train, test)
```



##Doing random forest analysis
```{r, warning = FALSE}
library(randomForest)
require(caTools)
library(caret)
```



```{r, warning = FALSE}
trainrf <- Survived~ Pclass+Sex+Age+SibSp+Parch+Fare
rf <- randomForest(trainrf, data = train, mtry = 3)
```

```{r}
varImpPlot(rf)
```
From this graph, it is clear that Sex is the best predictor variable of Survival of all these variables, followed by Fare and Age

#Creating Output Dataframe
```{r}
testrf <- "Pclass+Sex+Age+SibSp+Parch+Fare"
```

```{r}
Survivedpredict<- predict(rf, newdata = test)

```
```{r}
PassengerID <- test$PassengerId
outputdata <- as.data.frame(PassengerID)
outputdata$Survived <- Survivedpredict
```

```{r}
write.csv(outputdata, file = "A03_data_output.csv", row.names = FALSE)
```


