---
title: 'A03 Logistic regression '
author: "Cadence Fisher"
date: '2022-04-02'
slug: a03-logistic-regression
categories: []
tags: []
---
Use the Titanic passenger data (name, age, price of ticket, etc) to try to predict who will survive and who will die

your submission should have:

a "PassengerId" column containing the IDs of each passenger from test.csv.
a "Survived" column (that you will create!) with a "1" for the rows where you think the passenger survived, and a "0" where you predict that the passenger died.
#Loading in / Preparing the data

```{r, warning = FALSE}
library(openintro)
library(tidyverse)
library(ggridges)
library(tidymodels)

```
```{r}
train <- read_csv("https://raw.githubusercontent.com/fisherca/website_experiment/main/train.csv")
attach(train)
glimpse(train)
```
```{r}
test <- read_csv("https://raw.githubusercontent.com/fisherca/website_experiment/main/test.csv")
attach(test)
glimpse(test)
```


# Experimenting with the Data
## Quick tables to explore patterns
This was not in the instructions, but I thought it would help me visualize the data
### Sex
```{r}
table(train$Survived, train$Sex)
```

```{r}
train %>%
  ggplot(aes(x = Sex, fill = Survived)) +
  geom_bar(position = "fill") +
  labs(
    x = 'Sex', 
    fill = "Survived", 
    y = NULL,
    title = 'Survived vs. Sex')
```
More Females survived than males, and a far higher proportion of females survived overall
## Passenger Class
```{r}
table(train$Survived, train$Pclass)
```
```{r}
136/(80+136)
87/(97+87)
119/(372+119)
```
62% of the First class passengers survived, wheras only 47% of second class and 24% of third class passengers survived
```{r}
train %>%
  ggplot(aes(x = Pclass, fill = Survived)) +
  geom_bar(position = "fill") +
  labs(
    x = 'Passenger Class', 
    fill = "Survived", 
    y = NULL,
    title = 'Survived vs. Passenger Class')
```
The higher class passengers were slightly more likely to survive than the lower class passengers
## Siblings and Spouses
```{r}
table(train$Survived, train$SibSp)
```
```{r}
210/(398+210)
(112+13+4+3+1)/(97+112+15+13+12+4+15+3+5+1+7)
```

34.5% of people who had no siblings or spouse survived, whereas 46% of people with at least one sibling or spouse survived
```{r}
train %>%
  ggplot(aes(x = SibSp, fill = Survived)) +
  geom_bar(position = "fill") +
  labs(
    x = 'Siblings and Spouses', 
    fill = "Survived", 
    y = NULL,
    title = 'Survived vs. Siblings and Spouses')
```
Having a smallernumber of siblings or spouses seems to increase liklihood of survival slightly
### Parents and Childern
```{r}
table(train$Survived, train$Parch)
```
```{r}
233/(445+233)
(65+40+3+1+1)/(65+40+3+1+1+53+40+2+4+4+1)
```

Only 34% of people with no parents or children survived, but 51% of people with at least 1 parent or child survived
```{r}
train %>%
  ggplot(aes(x = Parch, fill = Survived)) +
  geom_bar(position = "fill") +
  labs(
    x = 'Parents and Childern', 
    fill = "Survived", 
    y = NULL,
    title = 'Survived vs. Parents and Childern')
```
People with a low to moderate amount of parents and children seem to be the most likely to survive
##Age
```{r}
train %>%
  ggplot(aes(x = Age, fill = Survived)) +
  geom_histogram(position = "fill", bins = 20) +
  labs(
    x = 'Age', 
    fill = "Survived", 
    y = NULL,
    title = 'Survived vs. Age')
```
Age does not seem to strongly predict survival, except very young and very old people seem slightly more likely to survive
## Fare
```{r}
train %>%
  ggplot(aes(x = Fare, fill = Survived)) +
  geom_histogram(position = "fill", bins = 10) +
  labs(
    x = 'Fare', 
    fill = "Survived", 
    y = NULL,
    title = 'Survived vs. Fare')
```
It seems like the more expensive the fare was, the more likely it was that the passenger would survive
## Is gender a good predictor of death? Recreation of exercise done in tutorial
```{r}
women_tot <- train %>%
  filter(Sex == 'female')
women_alive <- train %>%
  filter(Sex == 'female', Survived == 1)
women_rate = count(women_alive)/count(women_tot)
women_rate
```

74% of the women in the train data survived
```{r}
men_tot <- train %>%
  filter(Sex == 'male')
men_alive <- train %>%
  filter(Sex == 'male', Survived == 1)
men_rate = count(men_alive)/count(men_tot)
men_rate
```
18% of the men in the train data survived

#Building a model:
We are going to look at Pclass,Age, Fare Sex, SibSp, and Parch as potential predictors for Survived
##Preparing the Data
adding empty row to test data
```{r}
test$Survived <- NA
```
removing rows I will not use
```{r}
 
train <- train[c("Survived", "Pclass", "Sex", "Age", "SibSp", "Parch", "Fare")]

test <- test[c("Survived", "Pclass", "Sex", "Age", "SibSp", "Parch", "Fare")]
```

replacing NA values and changing data types if needed 
```{r}
train[train == ""] <- NA
train[is.na(train$Age), "Age"] <- median(train$Age, na.rm = TRUE)

test[is.na(test$Age), "Age"] <- median(test$Age, na.rm = TRUE)
```
```{r}
train[is.na(train$Fare), "Fare"] <- median(train$Fare, na.rm = TRUE)

test[is.na(test$Fare), "Fare"] <- median(test$Fare, na.rm = TRUE)
```
```{r}
train <- transform(train, Survived = as.factor(Survived), Sex = as.factor(Sex))

test <- transform(test, Sex = as.factor(Sex))
```

making a full dataset
```{r}
full_data <- rbind(train, test)
```

# Logistic Regression Practice
## Noting General Correlation Trends:
```{r}
train.numeric <- transform(train, Survived = as.numeric(Survived), Pclass = as.numeric(Pclass), Sex = as.numeric(Sex), Age = as.numeric(Age), SibSp = as.numeric(SibSp), Parch = as.numeric(Parch), Fare = as.numeric(Fare))
cor(train.numeric)
```

## Pclass
```{r}
logpclass <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(Survived~Pclass, data = train, family = "binomial")

tidy(logpclass)
```

```{r}
yhat <- predict(logpclass, test, type = "raw")
phat <- exp(yhat)/(1 + exp(yhat))
test <- test %>%
  bind_cols(yhat = yhat, phat = phat)

logpclassaug <- augment(logpclass$fit) %>%
  mutate(prob = exp(.fitted)/(1+exp(.fitted)))
logpclassaug
```
```{r}
ggplot(logpclassaug, aes(x = Pclass)) +
  geom_point(aes(y = as.numeric(Survived), color = Survived), alpha = 0.5) +
  scale_y_continuous(breaks = c(0,1)) +
  guides(color = "none") +
  geom_line(aes(y = prob)) +
  geom_point(data = test, aes(x = Pclass, y = phat),
             stroke = 1, size = 6) +
  labs(
    x = "Pclass",
    y = "Survived", 
    title = "Survived vs. Pclass"
  )
```

## Sex
```{r}
logsex <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(Survived~Sex, data = train, family = "binomial")

tidy(logsex)
```
```{r}
yhat2 <- predict(logsex, test, type = "raw")
phat2 <- exp(yhat2)/(1 + exp(yhat2))
test <- test %>%
  bind_cols(yhat2 = yhat2, phat2 = phat2)

logsexaug <- augment(logsex$fit) %>%
  mutate(prob2 = exp(.fitted)/(1+exp(.fitted)))
```
```{r}
ggplot(logsexaug, aes(x = Sex)) +
  geom_point(aes(y = as.numeric(Survived), color = Survived), alpha = 0.5) +
  scale_y_continuous(breaks = c(0,1)) +
  guides(color = "none") +
  geom_line(aes(y = prob2)) +
  geom_point(data = test, aes(x = Sex, y = phat2),
             stroke = 1, size = 6) +
  labs(
    x = "Sex",
    y = "Survived", 
    title = "Survived vs. Sex"
  )
```
## Age
```{r}
logage <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(Survived~Age, data = train, family = "binomial")

tidy(logage)
```
```{r}
yhat3 <- predict(logage, test, type = "raw")
phat3 <- exp(yhat3)/(1 + exp(yhat3))
test <- test %>%
  bind_cols(yhat3 = yhat3, phat3 = phat3)

logageaug <- augment(logage$fit) %>%
  mutate(prob3 = exp(.fitted)/(1+exp(.fitted)))
```
```{r}
ggplot(logageaug, aes(x = Age)) +
  geom_point(aes(y = as.numeric(Survived), color = Survived), alpha = 0.5) +
  scale_y_continuous(breaks = c(0,1)) +
  guides(color = "none") +
  geom_line(aes(y = prob3)) +
  geom_point(data = test, aes(x = Age, y = phat3),
             stroke = 1, size = 6) +
  labs(
    x = "Age",
    y = "Survived", 
    title = "Survived vs. Age"
  )
```
## SibSp
```{r}
logsibsp <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(Survived~SibSp, data = train, family = "binomial")

tidy(logsibsp)
```
```{r}
yhat4 <- predict(logsibsp, test, type = "raw")
phat4 <- exp(yhat4)/(1 + exp(yhat4))
test <- test %>%
  bind_cols(yhat4 = yhat4, phat4 = phat4)

logsibspaug <- augment(logsibsp$fit) %>%
  mutate(prob4 = exp(.fitted)/(1+exp(.fitted)))
```
```{r}
ggplot(logsibspaug, aes(x = SibSp)) +
  geom_point(aes(y = as.numeric(Survived), color = Survived), alpha = 0.5) +
  scale_y_continuous(breaks = c(0,1)) +
  guides(color = "none") +
  geom_line(aes(y = prob4)) +
  geom_point(data = test, aes(x = SibSp, y = phat4),
             stroke = 1, size = 6) +
  labs(
    x = "Siblings and Spouses",
    y = "Survived", 
    title = "Survived vs. Number of Siblings and Spouses"
  )
```

## Parch
```{r}
logparch <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(Survived~Parch, data = train, family = "binomial")

tidy(logparch)
```
```{r}
yhat5 <- predict(logparch, test, type = "raw")
phat5 <- exp(yhat5)/(1 + exp(yhat5))
test <- test %>%
  bind_cols(yhat5 = yhat5, phat5 = phat5)

logparchaug <- augment(logparch$fit) %>%
  mutate(prob5 = exp(.fitted)/(1+exp(.fitted)))
```
```{r}
ggplot(logparchaug, aes(x = Parch)) +
  geom_point(aes(y = as.numeric(Survived), color = Survived), alpha = 0.5) +
  scale_y_continuous(breaks = c(0,1)) +
  guides(color = "none") +
  geom_line(aes(y = prob5)) +
  geom_point(data = test, aes(x = Parch, y = phat5),
             stroke = 1, size = 6) +
  labs(
    x = "Parents and Children",
    y = "Survived", 
    title = "Survived vs. Number of Parents and Children"
  )
```

## Fare
```{r}
logfare <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(Survived~Fare, data = train, family = "binomial")

tidy(logfare)
```
```{r}
yhat6 <- predict(logfare, test, type = "raw")
phat6 <- exp(yhat6)/(1 + exp(yhat6))
test <- test %>%
  bind_cols(yhat6 = yhat6, phat6 = phat6)

logfareaug <- augment(logfare$fit) %>%
  mutate(prob6 = exp(.fitted)/(1+exp(.fitted)))

```
```{r}
ggplot(logfareaug, aes(x = Fare)) +
  geom_point(aes(y = as.numeric(Survived), color = Survived), alpha = 0.5) +
  scale_y_continuous(breaks = c(0,1)) +
  guides(color = "none") +
  geom_line(aes(y = prob6)) +
  geom_point(data = test, aes(x = Fare, y = phat6),
             stroke = 1, size = 6) +
  labs(
    x = "Fare",
    y = "Survived", 
    title = "Survived vs. Fare"
  )
```
