---
title: "Palmer Penguins Visualizations"
author: "Cadence Fisher"
date: '2022-02-09'
slug: palmer-penguins-visualizations
categories: R
tags:
- Penguins
- GGPLOT2
---

#Palmer Penguins

Measurements for penguin species, island in **Palmer Archipelago**, size (flipper length, body mass, bill dimensions), and sex.
##Call Libraries
```{r Libraries, message=FALSE}
library(tidyverse)
library(palmerpenguins)
```
##Describe the Data
```{r}
glimpse(penguins)
```



##Scatter Plot 
Start with the penguins data frame, map bill depth to the x-axis and map bill length to the y-axis. Represent each observation with a point and map species to the colour of each point. Title the plot "Bill depth and length", add the subtitle "Dimensions for Adelie, Chinstrap, and Gentoo Penguins", label the x and y axes as "Bill depth (mm)" and "Bill length (mm)", respectively, label the legend "Species", and add a caption for the data source. Finally, use a discrete colour scale that is designed to be perceived by viewers with common forms of colour blindness.
```{r Scatter Plot, echo=FALSE, warning=FALSE}
ggplot(data = penguins, 
       mapping = aes(x = bill_depth_mm, y = bill_length_mm,
                     colour = species, shape = island, size = body_mass_g, alpha = flipper_length_mm)) +
  geom_point(size = 2, alpha = 0.5) +
  labs(title = "Bill depth and length",
       subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
       x = "Bill depth (mm)", y = "Bill length (mm)",
       colour = "Species",  caption = "Source: Palmer Station LTER / palmerpenguins package") +
  scale_colour_viridis_d()
```

##Scatter Plot with Facet Grid
Useful for exploring conditional relationships and large data.
```{r Facet Plot, echo=FALSE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm)) + 
  geom_point() +
  facet_grid(species ~ island)
```

##Visualizing Categorical Data
```{r}
library(openintro)
loans <- loans_full_schema %>%
  select(loan_amount, interest_rate, term, grade, 
         state, annual_income, homeownership, debt_to_income)
glimpse(loans)
```

##Bar Plot
```{r}
ggplot(loans, aes(x = homeownership)) +
  geom_bar()
```

##Segmented Bar Chart
```{r}
ggplot(loans, aes(x = homeownership, 
                  fill = grade)) +
  geom_bar()
```
##Segmented Bar Chart Flipped
```{r}
ggplot(loans, aes(y = homeownership,
                  fill = grade)) +
  geom_bar(position = "fill") +
  labs(
    x = "Proportion",
    y = "Homeownership",
    fill = "Grade",
    title = "Grades of Lending Club loans",
    subtitle = "and homeownership of lendee"
  )
```

##Ridge Plots:
```{r}
library(ggridges)
ggplot(loans, aes(x = loan_amount, y = grade, fill = grade, color = grade)) + 
  geom_density_ridges(alpha = 0.5)
```

