---
title: Practice With Single Data Frame
author: Cadence Fisher
date: '2022-02-14'
slug: practice-with-single-data-frame
categories: []
tags: []
---
##Inserting Data Frame
```{r}
library(tidyverse)
url <- 'https://raw.githubusercontent.com/fisherca/tidyverse/main/hotels.csv'
hotels <- read_csv(url)
```
##Select to Exclude Variables
```{r}
hotels %>%
  select(-agent)
```
##Select a Range of Variables
```{r}
hotels %>%
  select(hotel:arrival_date_month)
```
##Select Variables With Certain Characteristics
```{r}
hotels %>%
  select(starts_with("arrival"))
```
##Arrange in Ascending vs Descending Order
```{r}
hotels %>%
  select(adults, children, babies) %>%
  arrange(babies)

hotels %>%
  select(adults, children, babies) %>%
  arrange(desc(babies))
```
## Slice for certain rows
```{r}
hotels %>%
  slice(1:5)
```
##Filter to select a subset of rows
```{r}
hotels %>%
  filter(hotel == "City Hotel")
```
##Filtering for many conditions at once
```{r}
hotels %>%
  filter( 
    adults == 0,
    children >= 1
    ) %>% 
  select(adults, babies, children)
```
##Filtering for complex conditions
```{r}
# bookings with no adults and some children or babies in the room
hotels %>%
  filter( 
    adults == 0,     
    children >= 1 | babies >= 1     # | means or
    ) %>%
  select(adults, babies, children)
```
##Using "distinct" to filter for unique rows
```{r}
hotels %>% 
  distinct(hotel, market_segment) %>%
  arrange(hotel, market_segment)
```
##Use count to create frequency tables
```{r}
# alphabetical order by default
hotels %>%
  count(market_segment)
# descending frequency order
hotels %>%
  count(market_segment, sort = TRUE)
```
##Using count and arrange
```{r}
# ascending frequency order
hotels %>%
  count(market_segment) %>%
  arrange(n)
# descending frequency order
# just like adding sort = TRUE
hotels %>%
  count(market_segment) %>%
  arrange(desc(n))
```
##Count for multivariate
```{r}
hotels %>%
  #order matters!
  count(hotel, market_segment)
```
#Mutate to add new variable
```{r}
hotels %>%
  mutate(little_ones = children + babies) %>%
  select(children, babies, little_ones) %>%
  arrange(desc(little_ones))
```
##Summary stats
```{r}
hotels %>%
  summarise(
    min_adr = min(adr),
    mean_adr = mean(adr),
    median_adr = median(adr),
    max_adr = max(adr)
    )
```
##Group by: for grouped operations
```{r}
# mean average daily rate for all booking at city and resort hotels
hotels %>%
  group_by(hotel) %>%
  summarise(mean_adr = mean(adr))
```

