---
title: "Data IO"
author: "Cadence Fisher"
date: '2022-02-23'
slug: data-io
categories: []
tags: []
---
# Reading rectangular dat into R
```{r}
library(tidyverse)
```

```{r}
# Will use packages readr and readxl
#readr:
#read_csv() - comma delimited files
#read_csv2() - semicolon separated files (common in countries where , is used as the decimal place)
#read_tsv() - tab delimited files
#read_delim() - reads in files with any delimiter
#read_fwf() - fixed width files

#readxl
#read_excel() - read xls or xlsx files
```

# Reading Data
```{r}
nobel <- read_csv("https://raw.githubusercontent.com/dearbharat/tidyverse/main/nobel.csv")
nobel
```

# Writing Data
```{r}
df <- tribble(
  ~x, ~y,
  1,  "a",
  2,  "b",
  3,  "c"
)
write_csv(df, file = "/Users/cfish/OneDrive/Documents/GitHub/website_experiment/content/post/2022-02-23-data-io/df.csv")
```
## Reading the file we just wrote:
```{r}
read_csv("/Users/cfish/OneDrive/Documents/GitHub/website_experiment/content/post/2022-02-23-data-io/df.csv")
```

# Practicing with Noble
```{r}
#already imported noble, but need to import sales too
sales <- read_csv("https://raw.githubusercontent.com/dearbharat/tidyverse/main/sales.csv")
```
## Split nobel into stem and non stem
```{r}
head(nobel)
```
```{r}
library(dplyr)
nobel %>%
  count(category)
```

```{r}
nobel_stem <- nobel %>%
  filter(category %in% "Physics" | category %in% "Medicine"| category %in% "Chemistry" | category %in% "Economics")
nobel_stem

```
```{r}
write_csv(nobel_stem, file = "/Users/cfish/OneDrive/Documents/GitHub/website_experiment/content/post/2022-02-23-data-io/noble_stem.csv")
```


```{r}
nobel_nonstem <- nobel %>%
  filter(category %in% "Literature" | category %in% "Peace")
nobel_nonstem
```
```{r}
write_csv(nobel_nonstem, file = "/Users/cfish/OneDrive/Documents/GitHub/website_experiment/content/post/2022-02-23-data-io/noble_nonstem.csv")
```

#Variable Names
## Data with bad names
```{r}
edibnb_badnames <- read_csv("https://raw.githubusercontent.com/dearbharat/tidyverse/main/edibnb-badnames.csv")
names(edibnb_badnames)
```
### R doesnt like spaces in variable names
```{r}
#ggplot(edibnb_badnames, aes(x = Number of bathrooms, y = Price)) +  geom_point()
#This will cause an error because of the names
```
## Option 1: define column names
```{r}
edibnb_col_names <- read_csv("https://raw.githubusercontent.com/dearbharat/tidyverse/main/edibnb-badnames.csv",
                             col_names = c("id", "price", 
                                           "neighbourhood", "accommodates",
                                           "bathroom", "bedroom", 
                                           "bed", "review_scores_rating", 
                                           "n_reviews", "url"))
names(edibnb_col_names)
```
## Option 2 - Format text to snake_case
```{r}
edibnb_clean_names <- read_csv("https://raw.githubusercontent.com/dearbharat/tidyverse/main/edibnb-badnames.csv") %>%
  janitor::clean_names()
names(edibnb_clean_names)
```

# Variable Types: if a data set is being read and one shows up as non applicable, how can we problem solve
## Option 1. Explicit NAs
```{r}
read_csv("https://raw.githubusercontent.com/dearbharat/tidyverse/main/df-na.csv", 
         na = c("", "NA", ".", "9999", "Not applicable"))
```
## Option 2. Specify column types
```{r}
read_csv("https://raw.githubusercontent.com/dearbharat/tidyverse/main/df-na.csv", col_types = list(col_double(), 
                                            col_character(), 
                                            col_character()))
```

# Case study: favourite foods:handle NAs, variable names, and make age numeric
```{r}
library(readxl)
```


fav_food <- read_excel("/Users/cfish/Downloads/Documents/favourite-food.xlsx",
                       na = c("N/A", "99999")) %>%
  janitor::clean_names()
fav_food <- fav_food %>%
  mutate(
    age = if_else(age == "five", "5", age),
    age = as.numeric(age)
    )
glimpse(fav_food)


