---
title: 'A02: Professor Diamond Data Regression Models'
author: "Cadence Fisher"
date: '2022-03-26'
slug: a02-professor-diamond-data-regression-models
categories: []
tags: []
---
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(psych)
```

```{r}
diamonds <- read_csv("https://raw.githubusercontent.com/fisherca/website_experiment/main/Professor_Proposes_Data.csv")
glimpse(diamonds)
```
```{r}
attach(diamonds)
```


```{r}
colnames(diamonds)
```
# Initial analysis of Data
## Price
Price has a bimodal distribution that is split at around $1,500. 
```{r}
ggplot(diamonds, aes(Price)) +
  geom_bar()
```
```{r}
describe(diamonds$Price)
```

## Removing Low Price Data
Because the Professor's diamond ended up being $3,000, we can conclude that he is not really interested in looking at diamonds in the lower hump of price's bimodal distribution
```{r}
diamonds_high <- diamonds %>%
  filter(diamonds$Price > 1500)
head(diamonds_high)
```

## Carat
From the graph below, we can gauge that carat for high price diamonds has a left skewed distribution
```{r}
ggplot(diamonds_high, aes(Carat)) +
  geom_bar()
```


```{r}
describe(diamonds_high$Carat)
```

## Colour
Colour has a unimodal, nearly symmetric distribution
```{r}
ggplot(diamonds_high, aes(Colour)) +
  geom_bar()
```
```{r}
diamonds_high %>%
  count(Colour)
```


## Clarity
clarity has a  distribution that is skewed a little to the left
```{r}
ggplot(diamonds_high, aes(Clarity)) +
  geom_bar()
```
```{r}
diamonds_high %>%
  count(Clarity)
```
### Making Clarity Numeric
```{r}
diamonds_high <- diamonds_high %>%
  mutate(Clarity = ifelse(Clarity %in% c("I1"), 1, ifelse(Clarity %in% c("I2"), 2, ifelse(Clarity %in% c("SI1"), 3, ifelse(Clarity %in% c("SI2"), 4, ifelse(Clarity %in% c("SI3"), 5, 6))))))
```


## Cut
Cut has a unimodal distribution skewed to the right
```{r}
ggplot(diamonds_high, aes(Cut)) +
  geom_bar()
```
```{r}
diamonds_high %>%
  count(Cut)
```
## Turning Cut into a numeric
```{r}
diamonds_high <- diamonds_high %>%
  mutate(Cut = ifelse(Cut %in% c("F"), 1, ifelse(Cut %in% c("G"), 2, ifelse(Cut %in% c("I"), 3, ifelse(Cut %in% c("V"), 4, 5)))))
```

## Certification
Certification has a unimodal distribution that is skewed to the right
```{r}
ggplot(diamonds_high, aes(Certification)) +
  geom_bar()
```
```{r}
diamonds_high %>%
  count(Certification)
```
```{r}

```
### Turning Certification Numeric
```{r}
diamonds_high <- diamonds_high %>%
  mutate(Certification = ifelse(Certification %in% c("AGS", "DOW"), 1, ifelse(Certification %in% c("EGL"), 2, 3)))
```

## Polish
polish has a bimodal distribution, split between F, G, and I; and v, V, and X
```{r}
ggplot(diamonds_high, aes(Polish)) +
  geom_bar()
```
```{r}
diamonds_high %>%
  count(Polish)
```
### Making Polish Numeric
```{r}
diamonds_high <- diamonds_high %>%
  mutate(Polish = ifelse(Polish %in% c("F", "G"), 1, ifelse(Polish %in% c("I", "v"), 2, 3)))
```


## Symmetry
symmetry has a bimodal distribution, with a low count of I splitting the bump of F and G with the bump of V and X
```{r}
ggplot(diamonds_high, aes(Symmetry)) +
  geom_bar()
```
```{r}
diamonds_high %>%
  count(Symmetry)
```
### Making Symmetry metric
```{r}
diamonds_high <- diamonds_high %>%
  mutate(Symmetry = ifelse(Symmetry %in% c("F", "G"), 1, ifelse(Symmetry %in% c("I", "V"), 2, 3)))
```
# Graphing the Data
## Carat

```{r}
ggplot(diamonds_high,aes(Carat, Price)) +
  geom_point(aes(color = Carat)) +
  geom_smooth(method='lm', se=FALSE, color='turquoise4') +
  theme_minimal() +
  labs(x='Carat', y='Price', title='Plot of Carat v.s. Price') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold')) 
```
```{r}
lmcarat <- lm(Price ~ Carat, data = diamonds_high)
summary(lmcarat)

```
```{r}
anovacarat <- aov(Price ~ Carat, data = diamonds_high)
summary(anovacarat)
```

The adjusted R-squared value of 0.1034 indicates that there is somewhat of a correlation between carat and price. Interestingly, for these higher priced diamonds, the higher carat diamonds are predicted to actually have a lower price
## Color
```{r}
lmcolor <- lm(Price ~ Colour, data = diamonds_high)
summary(lmcolor)
anovacolor <- aov(Price ~ Colour, data = diamonds_high)
summary(anovacolor)

```
With an adjusted R-squared of 0.06822, color seems like a pretty bad predictor of price


```{r}
ggplot(diamonds_high, aes(Colour, Price)) +
  geom_point() +
  theme_minimal() +
  labs(x='Colour', y='Price', title='Plot of Colour v.s. Price') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold')) 
```
There is basically no correlation between Colour and Price
## Clarity
```{r}
lmclarity <- lm(Price ~Clarity, data = diamonds_high)
summary(lmclarity)
anovaclarity <- aov(Price ~ Clarity, data = diamonds_high)
summary(anovaclarity)
```
With an adjusted R-squared of 0.1322,Clarity seems like an okay predictor of Price 

```{r}
ggplot(diamonds_high, aes(Clarity, Price)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE, color='red') +
  theme_minimal() +
  labs(x='Clarity', y='Price', title='Plot of Clarity v.s. Price') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold')) 
```
Generally, a higher clarity will cost more, but most of the midrange clarities are avaliable at many price ranges

# Cut
```{r}
lmcut <- lm(Price ~Cut, data = diamonds_high)
summary(lmcut)
anovacut <- aov(Price ~ Cut, data = diamonds_high)
summary(anovacut)
```
With an adjusted r squared of 0.0319, cut seems like it might be a pretty poor predictor of price
```{r}
ggplot(diamonds_high,aes(Cut, Price)) +
  geom_point(aes(color = Cut)) +
  geom_smooth(method='lm', se=FALSE, color='turquoise4') +
  theme_minimal() +
  labs(x='Cut', y='Price', title='Plot of Cut v.s. Price') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold')) 
```
Different levels of cuts can be found in a range of the higher prices
# Certification
```{r}
lmcert <- lm(Price ~Certification, data = diamonds_high)
summary(lmcert)
anovacert <- aov(Price ~ Certification, data = diamonds_high)
summary(anovacert)
```
With an adjusted Rsquared of 0.002633, Cetification seems like a terrible predictor of price
```{r}
ggplot(diamonds_high,aes(Certification, Price)) +
  geom_point(aes(color = Certification)) +
  geom_smooth(method='lm', se=FALSE, color='blue') +
  theme_minimal() +
  labs(x='Clarity', y='Price', title='Plot of Certification v.s. Price') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold')) 
```
# Polish
```{r}
lmpol <- lm(Price ~Polish, data = diamonds_high)
summary(lmpol)
anovapol <- aov(Price ~ Polish, data = diamonds_high)
summary(anovapol)
```
With an adjusted R-squared of 0.1003, Polish seems like a poor predictor of price
```{r}
ggplot(diamonds_high,aes(Polish, Price)) +
  geom_point(aes(color = Polish)) +
  geom_smooth(method='lm', se=FALSE, color='green') +
  theme_minimal() +
  labs(x='Polish', y='Price', title='Plot of Polish v.s. Price') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold')) 
```
# Symmetry
```{r}
lmsym <- lm(Price ~Symmetry, data = diamonds_high)
summary(lmsym)
anovasym <- aov(Price ~ Symmetry, data = diamonds_high)
summary(anovasym)
        
```
With an adjusted R-squared of 0.08865, symmetry seems like a poor predictor of price
```{r}
ggplot(diamonds_high,aes(Symmetry, Price)) +
  geom_point(aes(color = Symmetry)) +
  geom_smooth(method='lm', se=FALSE, color='turquoise4') +
  theme_minimal() +
  labs(x='Symmetry', y='Price', title='Plot of Symmetry v.s. Price') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold')) 
```
Generally, the more symmetric diamonds tend to cost a little bit more, but the professor can find a diamond of pretty good symmetry of a range of prices 


