---
title: "P01 Group Project"
author: "Cadence Fisher et al."
date: '2022-04-04'
slug: p01-group-project
categories: []
tags: []
---
Your task is to build a model to predict the churn of the bank customer by performing following steps:
The Main Question:
What factors affected wether the customer closed the bank account and left or continues to be a loyal customer?

# Perform Exploratory Data Analysis as necessary
```{r, warning=FALSE}
library(openintro)
library(tidyverse)
library(ggridges)
library(tidymodels)
library(data.table)
set.seed(1)

```
```{r}
Bank_Churn <- read_csv("https://raw.githubusercontent.com/fisherca/website_experiment/main/Bank_Churn.csv")
attach(Bank_Churn)
```
```{r}
head(Bank_Churn)
```
```{r}
summary(Bank_Churn)
```

## Check for NA and Blank values
```{r}
Bank_Churn <- Bank_Churn %>% mutate_all(na_if, "")
```

```{r}
isna <- apply(Bank_Churn,2, function(x) any(is.na(x)))
isna
```
None of the columns contain NA values or blank values
## Observation of the Variables
### Credit Score
```{r}
ggplot(Bank_Churn, aes(CreditScore)) + geom_bar(fill = "steelblue") +
  theme_minimal()
```

```{r}
ggplot(Bank_Churn, aes(x = CreditScore, fill = Exited)) +
  geom_bar(position = "fill") +
  labs(
    x = 'Credit Score', 
    fill = "Exited", 
    y = NULL,
    title = 'Exited vs. Credit Score') +
  scale_fill_manual(values = c("#DEB4A0", "#CA235F"))
```
### Geography:
```{r}
ggplot(Bank_Churn, aes(Geography)) + geom_bar(fill = "steelblue") +
  theme_minimal()
```
```{r}
Bank_Churn %>%
  ggplot(aes(x = Geography, fill = Exited)) +
  geom_bar(position = "fill") +
  labs(
    x = 'Geography', 
    fill = "Exited", 
    y = NULL,
    title = 'Exited vs. Geography') +
  scale_fill_manual(values = c("#DEB4A0", "#CA235F"))
```
### Gender
```{r}
ggplot(Bank_Churn, aes(Gender)) + geom_bar(fill = "steelblue") + theme_minimal()
```

```{r}
Bank_Churn %>%
  ggplot(aes(x = Gender, fill = Exited)) +
  geom_bar(position = "fill") +
  labs(
    x = 'Gender', 
    fill = "Exited", 
    y = NULL,
    title = 'Exited vs. Gender') +
  scale_fill_manual(values = c("#DEB4A0", "#CA235F"))
```
### Age
```{r}
ggplot(Bank_Churn, aes(Age)) + geom_bar(fill = "steelblue") +
  theme_minimal()
```
### Tenure (Number of years since opening bank account)
```{r}
ggplot(Bank_Churn, aes(Tenure)) + geom_bar(fill = "steelblue") +
  theme_minimal()
```
### Balance (current balance in bank account)
```{r}
ggplot(Bank_Churn, aes(Balance)) + geom_histogram(aes(fill = Balance, color = Balance), bins = 20) + theme_minimal()
```
### NumOfProducts 
```{r}
ggplot(Bank_Churn, aes(NumOfProducts)) + geom_bar()
```
### HasCrCard
```{r}
ggplot(Bank_Churn, aes(HasCrCard)) + geom_bar(fill = "steelblue") +
  theme_minimal()
```
### IsActiveMember
```{r}
ggplot(Bank_Churn, aes(IsActiveMember)) + geom_bar(fill = "steelblue") +
  theme_minimal()
```
### Estimated Salary
```{r}
ggplot(Bank_Churn, aes(EstimatedSalary)) + geom_histogram( aes(fill = EstimatedSalary), bins = 50) + theme_minimal()
```
### Exited: have they closed the bank account or stayed loyal. This is the best candidate for a dependent variable.
```{r}
ggplot(Bank_Churn, aes(Exited)) + geom_bar(fill = "steelblue") +
  theme_minimal()
```


#Balance the dataset before modeling (https://rpubs.com/DeclanStockdale/799284 (Links to an external site.)). Use any one of the over sampling, under sampling or SMOTE.

#Build a Logistic Regression model.

#Improve the model performance through transforming predictors (z-transform or log transform) or probability cut-off point of p.

#Evaluate and compare 2 - 3 different models.

#Explicitly write your assumptions and draw the conclusions.