---
title: "Practice with Data Types"
author: ''
date: '2022-02-21'
slug: practice-with-data-types
categories: []
tags: []
---
# First, install the readr package to be able to read csvs, and then load it using library(tidyverse)
```{r}
library(tidyverse)
```

# Load the Cat lovers data from github using its raw data's url
```{r}
cat_lovers <- read_csv("https://raw.githubusercontent.com/fisherca/tidyverse/main/cat-lovers.csv")
```

# Summarizing the Data
## A survey asked respondents their name and number of cats. The instructions said to enter the number of cats as a numerical value
## Try to take the mean:
```{r}
cat_lovers %>%
  summarise(mean_cats = mean(number_of_cats))
```
##Glimpse the data to see if there is a reason it is not working:
```{r}
glimpse(cat_lovers)
```

### The first assumption would be that maybe some did not enter a value, or entered "three" instead of "3" so enter na.rm that removes these values automatically:
```{r}
cat_lovers %>%
  summarise(mean_cats = mean(number_of_cats, na.rm = TRUE))
```
## Glimpse revealed that these "numeric" values are actually stored as character strings, so we need to convert the variables to numeric
```{r}
cat_lovers %>%
  mutate(number_of_cats = case_when(
    name == "Ginger Clark" ~ 2,
    name == "Doug Bass"    ~ 3,
    TRUE                   ~ as.numeric(number_of_cats)
    )) %>%
  summarise(mean_cats = mean(number_of_cats))
```
##Finally, make the function work so that ALL variables are converted to numeric

```{r}
cat_lovers %>%
  mutate(
    number_of_cats = case_when(
      name == "Ginger Clark" ~ "2",
      name == "Doug Bass"    ~ "3",
      TRUE                   ~ number_of_cats
      ),
    number_of_cats = as.numeric(number_of_cats)
    ) %>%
  summarise(mean_cats = mean(number_of_cats))
```

# Data Types
## Logical: TRUE and FALSE; boolean values (internally stored as 1 or 0)
```{r}
typeof(TRUE)
```
## Character: most flexible form
```{r}
typeof("hello")
```

##Double: floating point numerical values (the default type of numeric)
```{r}
#doubles are aka real numbers, but we call them doubles because the computer gives them double the space (a memory chain for the numbers before the decimal, and a memory chain for the numbers after the decimal). The precision you can achieve decreases the more numbers you have after the decimal
typeof(1.335)
typeof(7)
```

##Integer: 
```{r}
#integer numerical values (sometimes indicated with an L).
#byte is a collection of 8 binary digits (0s and 1s); there are 256 total distinct types of bytes. all data is stored in different multiples of bytes, typically 4 bytes per piece of data. The L for integer values indicates that the computer should "make space" for computations to be made with the integer, and give it double the space it needs. L aka "Large" integer
#signed integers (negative or specifically indicated to be positive) have slightly less capacity than unsigned integers. We have integers stored as signed in default
typeof(7L)
typeof(1:3)
```

##Info about how computers store things
```{r}
#A cluster is a collection of bytes (like 32 mb) that the computer reads, instead of reading data byte by byte. This allows the computer to be way faster
#r automatically assigns data to clusters, wheras its parent language c++ makes you do it manually
```
## Concatenation aka combining variables

##Explicit Coercion
```{r}
# Implicit vs explicit coercion: implicit is when the program automatically stores variables as one type, like if you have a data set with 1, 1.2, and 2, it will store the whole thing as a double, even though 2/3 are simple integers. Explicit is when we do this ourselves, like when we converted the cat data so we could analyze it
```

```{r}
#this is how you construct vectors, using the c() function
c(1, 2, 3)
c("Hello", "World!")
c(c("hi", "hello"), c("bye", "jello"))
```
```{r}
#convert with as.variabletype functions
x <- 1:3
typeof(x)
typeof(as.character(x))
y <- as.character(x)
y

a <- c(TRUE, FALSE)
a
b <- as.numeric(a)
b
```
```{r}
typeof(x)
typeof(y)
typeof(a)
typeof(b)
```

